The kinematics of the Local Group in a cosmological context

### 2013-02-08 paper/ms.tex
including new tables with a summary

\begin{abstract}

Recent observations constrained the tangential velocity of M31 with
respect to the Milky Way (MW) to be $v_{\rm M31,tan}<34.4$ \kms and
the radial velocity to be in the range $v_{\rm M31,rad}=-109\pm
4.4$\kms \citep{vanderMarel12}. In this study we use a large volume
high resolution N-body cosmological simulation to statistically study
this kinematics in the context of the $\Lambda$CDM cosmology. We
define the Local Group (LG) to be composed by two dominant separate
halos hosting the MW and the M31 galaxy fulfilling certain isolation
criteria. We find that the most probable values for the tangential and
radial velocities in these pairs are $v_{\rm rad, \Lambda CDM}=-60\pm
15$\kms and $v_{\rm tan, \Lambda CDM}=50\pm 5$\kms. Within a similar
absolute uncertainty defined by observations the pairs centered around
these values are $\sim3$ times more abundant than the pairs in the
observational interval. Furthermore, we find that only $\sim12\%$ of
the pairs show a ratio between its tangential and radial velocity
consistent with observations. This tension increases if we make a
narrower selection in the pair sample to match the observational
constraints on pair separation and total mass. Expressing these
results in terms of the angular momentum and mechanical energy per
unit mass also reflects this disagreement albeit in a less strong
fashion due to the larger uncertainties inferred for these physical
quantities. From these quantities Peeble's spin parameter is evaluated
and is used to characterize the LG with repect to other halo pairs in
$\Lambda$CDM. We also include in our analysis results from constrained
simulations. The characterisation results of this study suggest that
the formation and evolution of the LG differ from the average
$\Lambda$CDM halo pair. 
\end{abstract}

@@

Note that only one constrained LG-like object is included in the subsample.

\begin{table}
\caption{Summary of the kinematic properties for M31 in the galactocentric frame as reported by \citep{vanderMarel12}. Values in parenthesis correspond to vector components. $\sigma_{\bf x}$ represents the uncertainty on the components of vector ${\bf x}$}
\begin{center}
\begin{tabular}{ccc}\hline\hline
${\bf r}_{\rm M31}$ & (kpc) &$(-378.9, 612.7, -283.1)$\\
$\sigma_{{\bf r}, {\rm M31}}$ & (kpc) &$(-18.9, 30.6, 14.5)$\\
${\bf v}_{\rm M31}$ & (\kms) & $(66.1, -76.3, 45.1)$\\
$\sigma_{{\bf v},{\rm M31}}$ & (\kms) &$(26.7, 19.0, 26.5)$\\
$v_{\rm M31,rad}$ &(\kms) & $-109.3\pm 4.4$\\
$v_{\rm M31,tan}$ &(\kms) & $<34.4$\\
$r_{\rm M31}$ &(kpc) & $770\pm 40$\\
$M_{\rm 200, MW}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
$M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
$M_{\rm 200,MW} + M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $3.14\pm 0.58$\\\hline

\end{tabular}\\
\vspace{1mm}
Note: the observational uncertainties in the position vector
correspond to a $5\%$ in each component consistent with the
uncertainties in the distance \citep[see references
  in][]{vanderMarel08} 
\end{center}
\label{table:1}
\end{table}

@@

>>>>

\begin{abstract}
Recent observations constrained the tangential velocity of M31 with
respect to the Milky Way (MW) to be $v_{\rm M31,tan}<34.4$ \kms and
the radial velocity to be in the range $v_{\rm M31,rad}=-109\pm
4.4$\kms \citep{vanderMarel12}. In this study we use a large volume
high resolution N-body cosmologicalsimulation (BOLSHOI) to
statistically study this kinematics in the context of the $\Lambda$CDM
cosmology and three constrained simulations designed to reproduce the
Local Group (LG) and the local cosmic neighborhood. A LG is defined
here as  composed by two dominant separate halos hosting the MW and
the M31 galaxy fulfilling certain isolation criteria. An ensemble  of
all BOLSHOI pairs fulfilling this selection is constructed, and also a
subsample of it which obeys  more stringent observational constraints
on the mass and separation of of the MW and M31 (hereafter the full-
and sub- samples). The comparison of the ensembles of simulated pairs
with the observed LG has been done with respect to the observed radial
and tangential,  the reduced orbital energy ($e$) and angular momentum
($l$) of the LG, and with respect to the dimensionless spin parameter,
$\lambda$. Our main results are: 
%
%
% Jaime here I suggest to give in a very brief way the numbers of Table 2 that I have suggested.
%
The three constrained LGs are more consistent with the unconstrained
sample of LGs rather then with the observed LG.

It follows that the quasi-conserved quantities quantities that
characterize the orbit of the LG, i.e. $e$, $l$ and $\lambda$, do not
challenge the standard $\Lambda$CDM model, but that model is in
tension with regard to the particular phase on the orbit of the LG,
namely the actual values of the radial and tangential velocities. This
might hint to a problem of the $\Lambda$ CDM model to reproduce the
observed LG.

\end{abstract}

@@

Including these conditions the full pair sample is reduced from $1923$
to $158$ pairs. Note that only one constrained LG-like object is
included in the subsample.

\begin{table}
\caption{Summary of observational constraints. The kinematic
  properties for M31 are reported in the galactocentric
  \citep{vanderMarel12}. Values in parenthesis correspond to vector
  components. $\sigma_{\bf x}$ represents the uncertainty on the
  components of vector ${\bf x}$. The values for the individual halo
masses are consistent with the priors used by \cite{vanderMarel12}.}
\begin{center}
\begin{tabular}{ccc}\hline
$v_{\rm M31,rad}$ &(\kms) & $-109.3\pm 4.4$\\
$v_{\rm M31,tan}$ &(\kms) & $<34.4$\\
$r_{\rm M31}$ &(kpc) & $770\pm 40$\\
${\bf r}_{\rm M31}$ & (kpc) &$(-378.9, 612.7, -283.1)$\\
$\sigma_{{\bf r}, {\rm M31}}$ & (kpc) &$(-18.9, 30.6, 14.5)$\\
${\bf v}_{\rm M31}$ & (\kms) & $(66.1, -76.3, 45.1)$\\
$\sigma_{{\bf v},{\rm M31}}$ & (\kms) &$(26.7, 19.0, 26.5)$\\
$M_{\rm 200, MW}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
$M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
$M_{\rm 200,MW} + M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $3.14\pm 0.58$\\
$\log_{10}\lambda$& & $-1.72\pm 0.07$ \\\hline
\end{tabular}\\
\end{center}
\vspace{1mm}
Note: the observational uncertainties in the position vector
correspond to a $5\%$ in each component consistent with the
uncertainties in the distance \citep[see references
  in][]{vanderMarel08}.
Note: the value for $\log_{10}\lambda$ is obtained in this Letter
from a Monte Carlo simulation as described in Section XX.
\label{table:1}
\end{table}

\begin{table}
\caption{Summary of results from $\Lambda$CDM.}
\begin{center}
\begin{tabular}{cccc}\hline
 &  & Full Sample & Reduced Sample\\
$v_{\rm M31,rad}$ &(\kms) & $-70\pm 10$ & $-90\pm 10$\\
$v_{\rm M31,tan}$ &(\kms) & $50\pm 10$ & $50\pm 10$\\
$\log_{10}\lambda$& & $-1.47\pm 0.13$& \\ 
\end{tabular}
\end{center}
\vspace{1mm}
Note: the observational uncertainties in the position vector
correspond to a $5\%$ in each component consistent with the
uncertainties in the distance \citep[see references
  in][]{vanderMarel08}

\label{table:2}
\end{table}

@@

\begin{table*}
\caption{Summary of the comparison against $\Lambda$CDM.}
\begin{center}
\begin{tabular}{ccccc}\hline
Physical & (\%) Pairs consistent & \% Pairs consistent & \% Pairs
consistent& (\%) Pairs consistent\\ 
& with observations & with observations & with LCDM & with LCDM\\ 
 & in full sample & in reduced sample & in full sample & in reduced sample\\ \hline
$v_{\rm r}$-$v_{\rm t}$ & (0.4\%) 8/1923 & (0\%)0/158 & (1\%)23/1923 & (8\%)13/158\\
$e_{\rm tot}$-$l_{\rm orb}$ & (15\%)298/1923 & (24\%)38/158 & - & -\\
$\log_{10}\lambda$ & (13\%)257/1923 & (9\%)15/158 & - & -\\
$r_{\rm t}=v_{\rm t}/v_{\rm r}$& (12\%)242/1923 & (8\%)13/158 & -& -\\
\end{tabular}
\end{center}
\vspace{1mm}
Note: the observational uncertainties in the position vector
correspond to a $5\%$ in each component consistent with the
uncertainties in the distance \citep[see references
  in][]{vanderMarel08}

\label{table:2}
\end{table*}
